---
title: "01explore-and-check-assumptions"
format: html
---

```{r}
library(ggplot2) #creating plots
library(tidyverse) #data cleaning 
library(here) #managing directories, points to Rproj as working directory
library(DataExplorer) #probably won't use, auto generates exploratory analysis
library(gt) #for creating tables

all_data<- read_csv(here("data","clean_data.csv"),show_col_types = FALSE)
between_subj <- read_csv(here("data","between_subj.csv"), show_col_types = FALSE)
```

## Descriptive Statistics

We'll want to create a table with some basic descriptive statics of things like mean and standard deviations of vividness reports, accuracy in probe task, VVIQ scores. You can ignore this for the moment if you'd like or feel free to give it a go. I'd recommend starting with creating the graphs that I've suggested below.

## Univariate 

Explore each variable on its own.

-   Create some histogram (using geom_hist or geom_col) to see the distribution of of subjective responses, VVIQ scores, and vividness responses.

-   For across subjects use the all_data dataframe (trial-by-trial vividness, subjective responses). For between subjects use the between_subj dataframe (VVIQ scores, average vividness responses, congruence effect) . Use facet_wrap() to create separate plots for each participant.

## Bivariate

-   Explore the relationship between two variables using scatter plots (geom_point)

    -   Avg Vividness Responses x VVIQ(2)

    -   Subjective Priming Score x VVIQ(2)

    -   Subjective Priming Score x Avg Vividness Response

    -   Objective Priming Score x Avg Vividness Response (see example graph below that does something similar with VVIQ)

#### Objective Priming Score x VVIQ(2)

For the probe detection task :$$ Congruence Effect = CongruentTrialAccuracy - IncongruentAccuracy $$

```{r}
ggplot(between_subj, aes(VVIQ2,congruence_effect))+
  geom_point() + geom_hline(yintercept = 0)+ #adds horizontal line at 0
  geom_segment(aes(xend = VVIQ2, yend = 0)) #adds line connecting to each point to 0
```

#### Probe Accuracy

```{r}
plot_data <- clean_data %>%
  mutate(
    trial_type_clean = case_when(
      trial_type == "control" ~ "Control",
      trial_type == "congruent" ~ "Probe = Imagined Color",
      trial_type == "incongruent" ~ "Probe \u2260 Imagined Color",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(trial_type_clean)) %>%
  group_by(participant, trial_type_clean) %>%
  summarise(
    prop_correct = mean(correct, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    trial_type_clean = factor(
      trial_type_clean,
      levels = c(
        "Control",
        "Probe = Imagined Color",
        "Probe \u2260 Imagined Color"
      )
    )
  )

summary_data <- plot_data %>%
  group_by(trial_type_clean) %>%
  summarise(
    mean_acc = mean(prop_correct, na.rm = TRUE),
    se = sd(prop_correct, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

accuracy_plot <- ggplot(plot_data, aes(x = trial_type_clean, y = prop_correct)) +
  geom_line(aes(group = participant),
            color = "gray75",
            alpha = 0.3,
            linewidth = 0.7) +
  
  geom_point(aes(group = participant),
             color = "gray70",
             alpha = 0.45,
             size = 2) +
  
  geom_hline(yintercept = 0.5,
             linetype = "dashed") +
  
  geom_line(data = summary_data,
            aes(y = mean_acc, group = 1)) +
  
  geom_point(data = summary_data,
             aes(y = mean_acc)) +
  
  scale_y_continuous(
    limits = c(0, 1),
    breaks = c(0, 0.5, 1),
    labels = c("0%", "50%", "100%")
  ) +
  
  labs(
    title = "Accuracy by Trial Type",

    x = NULL,
    y = "Percent Correct"
  ) 
  
accuracy_plot
```
